Hunting experience shapes individual foraging specialization and predator-prey interactions in an online videogame

### 2023-04-18 manuscript/Chapter2-manuscript.Rmd
updated MS. full code for tables no to prevent bug

---
title: "Hunting experience shapes individual foraging specialization and predator-prey interactions in an online videogame"
output: 
  word_document:
    reference_docx: Chapter2-styles.docx
    fig_caption: yes
    #pandoc_args: ["-Fpandoc-crossref"] # to have eq. numbers
bibliography: references2.bib
csl: ecology.csl
---

@@

library(data.table)
library(brms)
library(flextable)
library(officer)
library(tidyr)

@@

dat[, c(2:6) := lapply(.SD, function (x) as.factor(x)), .SDcols = c(2:6)]

@@

ft <- compose(
  x = ft, j = 1,
  value = as_paragraph(
    as_equation(formula, height = .5)
  )
)

@@

ft <- my_theme(x = ft)
ft

@@

We parametrized every model described above to run four MCMC chains with 1000 posterior samples for each parameter. We ran 2500 iterations with a thinning set to eight for the additive model with a global smoother only (see Table I), and 1500 iterations with a thinning set to four for the additive model with a global smoother and group-level smoothers as well as for the additive model with group-level smoothers only (Table I). Burn-in was set to 500 iterations in each model. We parametrized the MDHGLM to run 2500 iterations with a thinning set to 8, with the first 500 iterations used as warmups. We assessed the convergence of the MCMC chains using trace plots, R-hat diagnostics with a threshold of <1.05, and effective sample sizes (ESS) with a threshold of >100 for the bulk-ESS and tail-ESS [@Vehtari.etal2021a]. We also performed posterior predictive checks which showed an adequate fit of the models (for details, see https://github.com/quantitative-ecologist/experience-hunting-tactics).

@@

tab1 <- readRDS(here::here("outputs", "05_outputs_tables", "05_table1.rds"))

tab1 <- autofit(tab1)

tab1 <- tab1 %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      "* 'elpd' refers to the 'expected log pointwise density' and is the value chosen to select the best model."),
    ref_symbols = " "
  ) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 11, part = "all")

set_caption(
  tab1,
  caption = "Leave-one-out cross-validation table of the three hierarchical GAMMs relating hunting success to player experience",
  style = "Table Caption"
)
```

@@

knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figure1.png"))

@@

tab2 <- readRDS(here::here("outputs", "05_outputs_tables", "05_table2.rds"))

tab2 <- autofit(tab2)

tab2 <- tab2 %>% 
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale. We back-transformed the hunting success values, estimated on a logit scale, back to a probability scale.",
      as_sup("\nb "),
      "The intercept values on the mean part of the equation for all traits indicate mean behaviour and success at the population level. The intercept values on the dispersion (i.e. sigma) part of the equation for predator speed indicate behavioural specialization at the population level."
    ),
    ref_symbols = " "
  ) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  width(j = c(2, 3, 4, 5), width = 3.7, unit = "cm") %>%
  width(j = 1, width = 3.2, unit = "cm")

set_caption(
  tab2,
  caption = "Posterior means and 95% credible intervals of the fixed effects estimated by the MDHGLM of predator speed, prey speed, and predator hunting success.",
  style = "Table Caption"
)

@@

tab3 <- readRDS(here::here("outputs", "05_outputs_tables", "05_table3.rds"))

tab3 <- autofit(tab3)

tab3 <- tab3 %>% 
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale. All the reported values are standard deviations.",
      as_sup("\nb "),
      "The intercept values on the mean part of the equation for all traits indicate among individual differences in mean behaviour and success.",
      as_sup("\nc "),
      "The intercept values on the dispersion (i.e. sigma) part of the equation for predator speed indicate among individual differences in behavioural specialization. For prey speed and hunting success, they indicate among individual differences in the variability of prey encounters and variability in hunting success, respectively."
    ),
    ref_symbols = " "
  ) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  width(j = c(2, 3, 4, 5), width = 3.7, unit = "cm") %>%
  width(j = 1, width = 3.2, unit = "cm")

set_caption(
  tab3,
  caption = "Posterior means and 95% credible intervals of the random effects estimated by the MDHGLM of predator speed, prey speed, and predator hunting success.",
  style = "Table Caption"
)

@@

knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figure2.png"))

@@

knitr::include_graphics(here::here("outputs", "05_outputs_figures", "05_figure3.png"))

>>>>

---
title: "Hunting experience shapes individual foraging specialization and predator-prey interactions in an online videogame"
output: 
  word_document:
    reference_docx: Chapter2-styles.docx
    fig_caption: yes
bibliography: references2.bib
csl: ecology.csl
---

@@

library(data.table)
library(brms)
library(flextable)
library(officer)
library(dplyr)

@@

dat[, c(2:6) := lapply(.SD, function(x) as.factor(x)), .SDcols = c(2:6)]

@@

# Transform values as a mathematical formula
ft <- compose(
  x = ft, j = 1,
  value = as_paragraph(
    as_equation(formula, height = .5)
  )
)

@@

# Apply theme and print
ft <- my_theme(x = ft)
ft

@@

We parametrized every model described above to run four MCMC chains with 1000 posterior samples for each parameter. We ran 2500 iterations with a thinning set to eight for the additive model with a global smoother only (see Table I), and 1500 iterations with a thinning set to four for the additive model with global and group-level smoothers and for the additive model with group-level smoothers only (Table I). We set the first 500 iterations in each model as warmups. We parametrized the MDHGLM to run 2500 iterations with a thinning set to 8, with the first 500 iterations used as warmups. We assessed the convergence of the MCMC chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 for the bulk-ESS and tail-ESS [@Vehtari.etal2021a]. We also performed posterior predictive checks which showed an adequate fit of the models (for details, see https://github.com/quantitative-ecologist/experience-hunting-tactics).

@@

# ===========================================================================
# 1. Import the model fit
# ===========================================================================

# Open the loo table
path <- file.path(here::here("outputs", "03_outputs_model-validation"))
loo_tab <- data.table(readRDS(file.path(path, "03A_lootable.rds")))

# Add model column
loo_tab[, model := c("Group-level smoothers only",
                     "Global smoother + group-level smoothers",
                     "Global smoother only")]

# Reorder columns
loo_tab <- loo_tab[, c(9, 1:8)]

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 2. Compute the table using flextable
# ===========================================================================

# Prepare the table parameters ----------------------------------------------

# Custom header
my_header <- data.frame(
   col_keys = c("model",
                "elpd_diff",
                "se_diff",
                "elpd_loo",
                "se_elpd_loo"),
   line1 = c("model",
             "elpd \ndifference",
             "sd \ndifference",
             "elpd loo \nvalue",
             "elpd loo \nstandard error"),
   stringsAsFactors = FALSE
)

# Custom theme
my_theme <- function(x, ...) {
   x <- colformat_double(x, big.mark = " ",
                         decimal.mark = ".",
                         digits = 2)
   x <- set_table_properties(x, layout = "fixed")
   x <- border_remove(x)
   std_border <- fp_border(width = 1, color = "black")
   x <- hline_top(x, part = "all", border = std_border)
   x <- hline_bottom(x, part = "all", border = std_border)
   autofit(x)
}

# Create the table ----------------------------------------------------------

tab1 <- loo_tab %>%
   select(model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
   flextable(col_keys = my_header$col_keys) %>%
   set_header_df(mapping = my_header, key = "col_keys") %>%
   my_theme() %>%
   merge_v(part = "header") %>%
   merge_h(part = "header") %>%
   fontsize(size = 10, part = "all") %>%
   font(fontname = "Times New Roman", part = "all") %>%
   align(align = "left", part = "all", j = 1) %>%
   align(align = "center", part = "all", j = 2) %>%
   align(align = "center", part = "all", j = 3) %>%
   align(align = "center", part = "all", j = 4) %>%
   align(align = "center", part = "all", j = 5)

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 3. Customize and print the table
# ===========================================================================

# Autofit to window
tab1 <- autofit(tab1)

# Add a footnote
tab1 <- tab1 %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      "* 'elpd' refers to the expected log pointwise density and is the value chosen to select the best model."),
    ref_symbols = " "
  ) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 11, part = "all")

# Table caption
set_caption(
  tab1,
  caption = "Leave-one-out cross-validation table of the three hierarchical GAMMs relating hunting success to player experience",
  style = "Table Caption"
)

# ===========================================================================
# ===========================================================================

@@

knitr::include_graphics(
  here::here("outputs", "05_outputs_figures", "05_figure1.png")
)

@@

# ===========================================================================
# 1. Import the model fit
# ===========================================================================

# path
path <- file.path(here::here("outputs"))

# Import the model
fit <- readRDS(
  file.path(path, "02_outputs_models", "02B_DHMLM.rds")
)

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 2. Fixed effects table
# ===========================================================================

# Prepare the table ---------------------------------------------------------

# Extract the fixed effects from the model
eff_tab <- data.table(fixef(fit), keep.rownames = TRUE)
eff_tab <- eff_tab[, c(1, 2, 4:5)]

# Rename variables
setnames(
  eff_tab,
  c("rn", "Q2.5", "Q97.5"),
  c("parameter", "lower_ci", "upper_ci")
)

# Add a column that specifies the experience level
eff_tab[parameter %like% "novice", xp_level := "novice"]
eff_tab[parameter %like% "interm", xp_level := "intermediate"]
eff_tab[parameter %like% "advanced", xp_level := "advanced"]

# Back transform success
eff_tab[
  parameter %like% "success",
  c(2:4) := lapply(.SD, function(x) {plogis(x)}),
  .SDcols = c(2:4)
]

# Back transform sigma
eff_tab[
  parameter %like% "sigma",
  c(2:4) := lapply(.SD, function(x) {exp(x)}),
  .SDcols = c(2:4)
]

# Round values to 3 digits
eff_tab[, c(2:4) := round(eff_tab[, c(2:4)], digits = 3)]

# Reorganize the table ------------------------------------------------------

# reorder columns, and then rows based on xp level
eff_tab <- eff_tab[, c(5, 1:4)]
setorder(
  eff_tab,
  cols = -"xp_level"
)

# Create trait column
eff_tab[
  , trait := rep(c(
    "predator speed", "predator speed",
    "prey speed", "prey speed", "hunting success"),
    6
  )
]

# Rename parameter
eff_tab[
   , parameter := rep(
      c(
         rep(c("intercept (mean)", "intercept (sigma)"), 2),
         "intercept (mean)",
         rep(c("prey rank (mean)", "prey rank (sigma)"), 2),
         "match duration (mean)"
      ),
      3
   )
]

# Paste upper and lower ci with estimate
eff_tab[, estimate := paste(format(Estimate, digits = 3), "(")]
eff_tab[, estimate := paste(estimate, format(lower_ci, digits = 3), sep = "")]
eff_tab[, estimate := paste(estimate, ",", sep = "")]
eff_tab[, estimate := paste(estimate, format(upper_ci, digits = 3), sep = " ")]
eff_tab[, estimate := paste(estimate, ")", sep = "")]
eff_tab[, c("Estimate", "lower_ci", "upper_ci") := NULL]

# Reshape
eff_tab <- dcast(
  eff_tab,
  trait + parameter ~ xp_level,
  value.var = "estimate"
)

# Reorder columns
eff_tab <- eff_tab[, c(1, 2, 5, 4, 3)]

# Reorder rows
table <- eff_tab[c(3, 5, 4, 6, 7, 9, 8, 10, 1, 2)]

# Because I can't align hunting success properly
table[10, 1] <- " "

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 3. Create the table using flextable
# ===========================================================================

# Prepare the table parameters ----------------------------------------------

# Custom header
my_header <- data.frame(
  col_keys = c("trait", "parameter",
               "novice", "intermediate", "advanced"),
  line1 = c("Trait",
            "Parameter",
            "Novice",
            "Intermediate",
            "Advanced"),
  stringsAsFactors = FALSE
)

# Custom theme
my_theme <- function(x, ...) {
  x <- colformat_double(x, big.mark = " ",
                        decimal.mark = ".",
                        digits = 3)
  x <- set_table_properties(x, layout = "fixed")
  x <- border_remove(x)
  std_border <- fp_border(width = 1, color = "black")
  x <- hline_top(x, part = "all", border = std_border)
  x <- hline_bottom(x, part = "all", border = std_border)
  autofit(x)
}

# Create the table ----------------------------------------------------------

tab2 <-

  # table structure
  table %>%
  select(trait, parameter,
         novice, intermediate, advanced) %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%
  # Align repeated rows on column 1
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%
  # font
  fontsize(size = 10, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%

  # Align text
  align(align = "left", part = "all", j = 1) %>%
  align(align = "left", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5)

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 4. Customize and print the table
# ===========================================================================

# Autofit to window size
tab2 <- autofit(tab2)

# Table footnote
tab2 <- tab2 %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale. We back-transformed the hunting success values, estimated on a logit scale, back to a probability scale.",
      as_sup("\nb "),
      "The intercept values on the mean part of the equation for all traits indicate mean behaviour and success at the population level. The intercept values on the dispersion (i.e. sigma) part of the equation for predator speed indicate behavioural specialization at the population level."
    ),
    ref_symbols = " "
  ) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  width(j = c(2, 3, 4, 5), width = 3.7, unit = "cm") %>%
  width(j = 1, width = 3.2, unit = "cm")

# Table caption
set_caption(
  tab2,
  caption = "Posterior means and 95% credible intervals of the fixed effects estimated by the MDHGLM of predator speed, prey speed, and predator hunting success.",
  style = "Table Caption"
)

# ===========================================================================
# ===========================================================================

@@

# ===========================================================================
# 1. Import the model fit
# ===========================================================================

# path
path <- file.path(here::here("outputs"))

# Import the model
fit <- readRDS(
  file.path(path, "02_outputs_models", "02B_DHMLM.rds")
)

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 2. Random effects table
# ===========================================================================

# Prepare data --------------------------------------------------------------

# Extract the random effect variances
sd_draws <- data.table(
   as_draws_df(
      fit, variable = "^sd_", regex = TRUE
   )
)

sd_draws[, c(28:30) := NULL]

# Calculate parameters -------------------------------------------------------

# Take the exponent of sigma parameters
sd_draws[
   , c("sd_predator_id__sigma_speednovice_Intercept",
       "sd_predator_id__sigma_speedinterm_Intercept",
       "sd_predator_id__sigma_speedadvanced_Intercept",
       "sd_predator_id__sigma_preyspeednovice_Intercept",
       "sd_predator_id__sigma_preyspeedinterm_Intercept",
       "sd_predator_id__sigma_preyspeedadvanced_Intercept"
   )
   := lapply(.SD, function(x) {exp(x)}),
   .SDcols = c(
      "sd_predator_id__sigma_speednovice_Intercept",
      "sd_predator_id__sigma_speedinterm_Intercept",
      "sd_predator_id__sigma_speedadvanced_Intercept",
      "sd_predator_id__sigma_preyspeednovice_Intercept",
      "sd_predator_id__sigma_preyspeedinterm_Intercept",
      "sd_predator_id__sigma_preyspeedadvanced_Intercept"
   )
]

# Reshape the table -----------------------------------------------------------

ranef_tab <- melt(
   sd_draws,
   variable.name = "parameter"
)

# Summarize the values (means + 95% CI) --------------------------------------

# Intervals
lower_interval <- function(x) {
  coda::HPDinterval(as.mcmc(x), 0.95)[1]
}
upper_interval <- function(x) {
  coda::HPDinterval(as.mcmc(x), 0.95)[2]
}

# Summarize values
ranef_tab[, ":=" (mean = mean(value),
                  lower_ci = lower_interval(value),
                  upper_ci = upper_interval(value)),
          by = .(parameter)]

ranef_tab <- unique(ranef_tab[, .(parameter, mean, lower_ci, upper_ci)])
ranef_tab[, c(2:4) := round(ranef_tab[, c(2:4)], 3)]

# Rearrange the table --------------------------------------------------------

# Add a column that specifies the experience level
ranef_tab[parameter %like% "novice", xp_level := "novice"]
ranef_tab[parameter %like% "interm", xp_level := "intermediate"]
ranef_tab[parameter %like% "advanced", xp_level := "advanced"]

# Paste upper and lower ci with estimate
ranef_tab[, estimate := paste(format(mean, digits = 3), "(")]
ranef_tab[, estimate := paste(estimate, format(lower_ci, digits = 3), sep = "")]
ranef_tab[, estimate := paste(estimate, ",", sep = "")]
ranef_tab[, estimate := paste(estimate, format(upper_ci, digits = 3), sep = " ")]
ranef_tab[, estimate := paste(estimate, ")", sep = "")]
ranef_tab[, c("mean", "lower_ci", "upper_ci") := NULL]

# Parameter column
ranef_tab[
   , parameter1 := ifelse(
       parameter %like% "predator_id__sigma",
       "predator ID (sigma)",
       "predator ID (mean)"
   )
]
ranef_tab[parameter %like% "avatar", parameter1 := "avatar (mean)"]
ranef_tab[parameter %like% "env", parameter1 := "environment (mean)"]

# Trait column
ranef_tab[
   , trait := ifelse(
       parameter %like% "preyspeed",
       "prey speed",
       "predator speed"
   )
]
ranef_tab[parameter %like% "success", trait := "hunting success"]

# Remove old parameter column and replace by new
ranef_tab[, parameter := NULL]
setnames(ranef_tab, "parameter1", "parameter")

# Reshape
ranef_tab <- dcast(
   ranef_tab,
   trait + parameter ~ xp_level,
   value.var = "estimate"
)

# Reorder columns
ranef_tab <- ranef_tab[, c(1, 2, 5, 4, 3)]

# Reorder rows
table <- ranef_tab[c(1:5, 6:9)]

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 3. Create the table using flextable
# ===========================================================================

# Prepare the table parameters ----------------------------------------------

# Custom header
my_header <- data.frame(
  col_keys = c("trait", "parameter",
               "novice", "intermediate", "advanced"),
  line1 = c("Trait",
            "Parameter",
            "Novice",
            "Intermediate",
            "Advanced"),
  stringsAsFactors = FALSE
)

# Custom theme
my_theme <- function(x, ...) {
  x <- colformat_double(x, big.mark = " ",
                        decimal.mark = ".",
                        digits = 3)
  x <- set_table_properties(x, layout = "fixed")
  x <- border_remove(x)
  std_border <- fp_border(width = 1, color = "black")
  x <- hline_top(x, part = "all", border = std_border)
  x <- hline_bottom(x, part = "all", border = std_border)
  autofit(x)
}

# Create the table ----------------------------------------------------------

tab3 <-

  # table structure
  table %>%
  select(trait, parameter,
         novice, intermediate, advanced) %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%

  # Align repeated rows on column 1
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%
  # font
  fontsize(size = 10, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%

  # Align text
  align(align = "left", part = "all", j = 1) %>%
  align(align = "left", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5)

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 4. Customize and print the table
# ===========================================================================

# Autofit to window
tab3 <- autofit(tab3)

# Add footnote
tab3 <- tab3 %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale. All the reported values are standard deviations.",
      as_sup("\nb "),
      "The intercept values on the mean part of the equation for all traits indicate among individual differences in mean behaviour and success.",
      as_sup("\nc "),
      "The intercept values on the dispersion (i.e. sigma) part of the equation for predator speed indicate among individual differences in behavioural specialization. For prey speed and hunting success, they indicate among individual differences in the variability of prey encounters and variability in hunting success, respectively."
    ),
    ref_symbols = " "
  ) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  width(j = c(2, 3, 4, 5), width = 3.7, unit = "cm") %>%
  width(j = 1, width = 3.2, unit = "cm")

# Add table caption
set_caption(
  tab3,
  caption = "Posterior means and 95% credible intervals of the random effects estimated by the MDHGLM of predator speed, prey speed, and predator hunting success.",
  style = "Table Caption"
)

# ===========================================================================
# ===========================================================================

@@

knitr::include_graphics(
  here::here("outputs", "05_outputs_figures", "05_figure2.png")
)

@@

knitr::include_graphics(
  here::here("outputs", "05_outputs_figures", "05_figure3.png")
)
```
